Article data analysis 2

### 2019-01-25 notebooks/article_notebook2.Rmd
minor changes, added figures to article notebooks

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, Biostrings, seqinr, stringr)
```

@@

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","oqx","qep","mfp","mcb")
mut_genes <- c("gyr","par","marR","marA","acrB","soxR","acrR","nor","ydh","sdi","rpoB")

@@

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

@@

## List of genes of interest
This table lists all the genes of interest in this study, some information about them, and their accession number(s).
```{r}

acquired_ref_names <- acquired_data$ref_name
acquired_ref <- c()

for (i in ac_genes) {
  for (j in acquired_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      acquired_ref <- c(acquired_ref, j)
    }
  }
}

acquired_ref <- unique(acquired_ref)

acquired_ref_genes <- acquired_data %>%
  mutate(test = duplicated(ref_name),
         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9",
                          grepl("qep", ref_name, ignore.case = T) == TRUE ~ "qepA")) %>%
  filter(test == FALSE,
         ref_name %in% acquired_ref_names) %>%
  select(ref_name, gene) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  dplyr::rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
         Type = "Acquired")

df_ref_names <- mut_data$cluster
ref_names <- c()

for (i in mut_genes) {
  for (j in df_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      ref_names <- c(ref_names, j)
    }
  }
}

ref_names <- unique(ref_names)

intrinsic_ref_genes <- mut_data %>%
  mutate(test = duplicated(cluster),
         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
                          grepl("soxR", cluster, ignore.case = T) == TRUE ~ "soxR",
                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
                          grepl("mara", cluster, ignore.case = T) == TRUE ~ "marA",
                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR",
                          grepl("acrb", ref_name, ignore.case = T) == TRUE ~ "acrB",
                          grepl("mdfa", ref_name, ignore.case = T) == TRUE ~ "mdfA",
                          grepl("rpob", ref_name, ignore.case = T) == TRUE ~ "rpoB")) %>%
  filter(test == FALSE,
         cluster %in% ref_names) %>%
  select(ref_name, gene, free_text) %>%
  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
  select(-ref_name) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  dplyr::rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
                                 Gene == "acrB" ~ "Gene that codes for the AcrB efflux pump, part of the AcrAB-TolC efflux system",
                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
                                 Gene == "marA" ~ "Transcriptional activator og genes involved in the MAR phenotype.",
                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
                                 Gene == "soxR" ~ "Transcriptional activator of the superoxide response regulon.",
                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump.",
                                 Gene == "mdfA" ~ "Chromosomal efflux pump.",
                                 Gene == "rpoB" ~ "Beta-subunit of RNA polymerase, mutations lead to increased expression of the ydhE-pump"),
         Type = "Intrinsic gene")

ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)

ref_genes

@@

@@

@@

@@

## Specific mutations in QRDR
```{r}
suppressWarnings(clean_mut_data %>%
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  select(-c(id, flag)) %>%
  gather(gene, mut, -ref) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  select(-mut) %>%
  dplyr::rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  group_by(gene, mut, nt_change) %>%
  dplyr::count() %>%
  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1)) %>%
  dplyr::rename("Mutation" = mut))
```

## Mutation combinations in QRDR
```{r}
mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  dplyr::count() %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1))
```

# Figures
## Resistance mechanisms per sequence type
```{r}
genes <- c(names(mut_report), names(acquired_report), "qnr")
genes <- genes[!genes %in% c("id","nt_change","qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")]

no_of_mlst <- mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(ST, species) %>%
  dplyr::count() %>%
  group_by(ST) %>%
  mutate(group_n = sum(n)) %>%
  filter(group_n > 4)

mlst_mechanisms_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,12:18]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mlst_data[c("id","ST")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, ST), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,ST)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  group_by(ST, gene, value) %>%
  dplyr::count() %>%
  filter(ST %in% no_of_mlst$ST)

>>>>

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable, ape, cluster, tibble, Biostrings, seqinr, stringr, scales)
```

@@

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","oqx","qep","aac6")
mut_genes <- c("gyr","par","marR","marA","sox","rpoB","rob","gad")

@@

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table)) %>%
  mutate(parC = ifelse(parC == "S58I", "S80I", parC)) %>%
  left_join(id_data, by = "ref") %>%
  select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S"))

@@

## List of genes of interest
This table lists all the genes of interest in this study, some information about them, and their accession number(s).
```{r}
# 
# acquired_ref_names <- acquired_data$ref_name
# acquired_ref <- c()
# 
# for (i in ac_genes) {
#   for (j in acquired_ref_names) {
#     if (grepl(i,j, ignore.case = T) == TRUE) {
#       acquired_ref <- c(acquired_ref, j)
#     }
#   }
# }
# 
# acquired_ref <- unique(acquired_ref)
# 
# acquired_ref_genes <- acquired_data %>%
#   mutate(test = duplicated(ref_name),
#          gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
#                           grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
#                           grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
#                           grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
#                           grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
#                           grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
#                           grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
#                           grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9",
#                           grepl("qep", ref_name, ignore.case = T) == TRUE ~ "qepA")) %>%
#   filter(test == FALSE,
#          ref_name %in% acquired_ref_names) %>%
#   select(ref_name, gene) %>%
#   mutate(compl = complete.cases(.)) %>%
#   filter(compl == TRUE) %>%
#   mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
#   group_by(gene) %>%
#   summarise_all(funs(func_paste)) %>%
#   select(gene, ref_name2) %>%
#   dplyr::rename("Accession number(s)" = ref_name2,
#          "Gene" = gene) %>%
#   mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
#          Type = "Acquired")
#   
# df_ref_names <- mut_data$cluster
# ref_names <- c()
# 
# for (i in mut_genes) {
#   for (j in df_ref_names) {
#     if (grepl(i,j, ignore.case = T) == TRUE) {
#       ref_names <- c(ref_names, j)
#     }
#   }
# }
# 
# ref_names <- unique(ref_names)
# 
# intrinsic_ref_genes <- mut_data %>%
#   mutate(test = duplicated(cluster),
#          gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
#                           grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
#                           grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
#                           grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
#                           grepl("soxS", cluster, ignore.case = T) == TRUE ~ "soxS",
#                           grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
#                           grepl("mara", cluster, ignore.case = T) == TRUE ~ "marA",
#                           grepl("rpob", ref_name, ignore.case = T) == TRUE ~ "rpoB",
#                           grepl("rob", ref_name, ignore.case = T) == TRUE ~ "robA")) %>%
#   filter(test == FALSE,
#          cluster %in% ref_names) %>%
#   select(ref_name, gene, free_text) %>%
#   mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
#   select(-ref_name) %>%
#   group_by(gene) %>%
#   summarise_all(funs(func_paste)) %>%
#   select(gene, ref_name2) %>%
#   dplyr::rename("Accession number(s)" = ref_name2,
#          "Gene" = gene)
#   # mutate(Description = case_when(Gene == "gyrA" ~ ,
#   #                                Gene == "gyrB" ~ ,
#   #                                Gene == "marR" ~ ,
#   #                                Gene == "marA" ~ ,
#   #                                Gene == "parC" ~ ,
#   #                                Gene == "parE" ~ ,
#   #                                Gene == "soxS" ~ "Transcriptional activator of superoxide response regulon",
#   #                                Gene == "rpoB" ~ "RNA polymerase subunit B",
#   #                                Gene == "robA" ~ ),
#   #        Type = "Intrinsic gene")
#                 
# ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)
# 
# ref_genes

@@

## Most abundant (n > 3) sequence types per animal species
```{r}
mlst_n <- mlst_data[,c("id","ST")] %>%
  group_by(ST) %>%
  dplyr::count() %>%
  filter(n > 3)

mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  spread(ST, species) %>%
  summarise_all(funs(func_paste)) %>%
  select(-id) %>%
  gather(ST, species) %>%
  filter(ST %in% mlst_n$ST)
```

@@

## Percent isolates with qnr family genes
```{r}
acquired_report %>%
  select(-c(qepA4, id)) %>%
  mutate_all(funs(as.numeric)) %>%
  dplyr::rename("qnrA" = qnrA1) %>%
  mutate(qnrB = ifelse(rowSums(.[,2:4]) >= 1, 1, 0),
         qnrS = ifelse(rowSums(.[,5:7]) >= 1, 1, 0)) %>%
  select(qnrA, qnrB, qnrS) %>%
  gather(gene, value) %>%
  group_by(gene, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 0, "Absent", "Present")) %>%
  spread(value, n) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1))
```

@@

## qnr-genes per animal species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")

acquired_report %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, qnr_genes) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  arrange(key, desc(Percent))
```

@@

## Percent mutations in each intrinsic gene
```{r}
mut_report %>%
  select(-nt_change) %>%
  gather(gene, value, -id) %>%
  group_by(gene, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1)) %>%
  arrange(desc(Percent))
```

## Specific mutations in QRDR
```{r}
suppressWarnings(clean_mut_data %>%
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  select(-c(id, flag)) %>%
  left_join(id_data, by = "ref") %>%
  dplyr::select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  gather(gene, mut, -id) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
  select(-mut) %>%
  dplyr::rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  group_by(gene, mut, nt_change) %>%
  dplyr::count() %>%
  separate(nt_change, into = c("Reference nt", "Contig nt"), sep = "-") %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1)) %>%
  dplyr::rename("Mutation" = mut) %>%
  arrange(gene, desc(Percent)))
```

## Specific QRDR mutations per animal species
```{r}
suppressWarnings(clean_mut_data %>%
  mutate(ref_ctg_change = ifelse(ref_ctg_change == ".", "", ref_ctg_change),
         ref_nt = ifelse(ref_nt == ".", "", ref_nt),
         ctg_nt = ifelse(ctg_nt == ".", "", ctg_nt),
         ref_ctg_nt_change = paste(ref_nt, ctg_nt, sep = "-"),
         mutation = paste(ref_ctg_change, ref_ctg_nt_change, sep = "/"),
         mutation = ifelse(mutation == "/-", "0", mutation)) %>%
  select(ref, gene_names, flag, mutation) %>%
  filter(flag %in% flag_selection,
         gene_names %in% c("gyrA","gyrB","parC","parE")) %>%
  mutate(id = 1:n()) %>%
  spread(gene_names, mutation) %>%
  group_by(ref) %>%
  summarise_all(funs(func_paste2)) %>%
  dplyr::select(-c(id, flag)) %>%
  left_join(id_data, by = "ref") %>%
  dplyr::select(id, everything(), -ref) %>%
  filter(id %not_in% c("2016-17-565-1-S",
                       "2016-02-428-2-S",
                       "2016-02-486-2-S",
                       "2016-02-732-2-S",
                       "2014-01-1741-1-S")) %>%
  gather(gene, mut, -id) %>%
  mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
         result = ifelse(mut != 0, 1, 0),
         result = as.integer(result),
         type = "mut") %>%
  separate(mut, into = paste("v", 1:12, sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:12, sep = "_")) %>%
  separate(value, into = c("aa_change","nt_change"), sep = "/") %>%
  mutate(aa_change = ifelse(gene == "parC" & aa_change == "S58I", "S80I", aa_change)) %>%
  select(-mut) %>%
  dplyr::rename("mut" = aa_change) %>%
  filter(is.na(nt_change) == FALSE) %>%
  fix_gyr_par_results(.) %>%
  filter(result_total == 1) %>%
  left_join(isolate_data[c("id","species")], by = "id") %>%
  group_by(gene, mut, species) %>%
  dplyr::count() %>%
  rowwise() %>%
  mutate(Total = case_when(species == "Broiler" ~ 87,
                           species == "Pig" ~ 75,
                           species == "Red fox" ~ 52,
                           species == "Wild bird" ~ 66),
         Percent = round(n/Total * 100,1)) %>%
  dplyr::rename("Mutation" = mut) %>%
  arrange(gene, desc(Percent)))
```

## Mutation combinations in QRDR
```{r}
mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  dplyr::count() %>%
  rowwise() %>%
  mutate(Total = 280,
         Percent = round(n/Total * 100,1)) %>%
  arrange(desc(Percent))
```

## Mutations in intrinsic genes per animal species
```{r}
gene_names <- names(mut_report)

mut_report %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, gene_names) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  dplyr::rename("Gene" = key,
                "Species" = species) %>%
  dplyr::select(Gene, Species, Present, Total, Percent) %>%
  arrange(Gene, desc(Percent))
```

## Mechanisms in isolates without QRDR mutations
```{r}
mut_report %>%
  filter(gyrA == 0,
         parC == 0,
         parE == 0) %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,14:20]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  select(-c(nt_change, id)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  dplyr::select(-c(gyrA,gyrB,parC,parE))
```

## Total combinations of resistance mechanisms
```{r}
mut_report %>%
  select(-c(nt_change, gyrA, parC, parE)) %>%
  left_join(mut_quant[c("gyrA","parC","parE","id")], by = "id") %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  arrange(desc(CIP))
```

## Chi Squared test for significant differences in resistance mechanisms
```{r}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    unlist() %>% unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

do_chisq_test <- function(df) {
  df <- df %>%
    dplyr::slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
          dplyr::filter(Var2 < Var1) %>% t) %>%
    mutate(test = rep(1:(n()/2), each = 2)) %>%
    group_by(test) %>%
    dplyr::do(data_frame(gene = .$key,
                  test = dplyr::first(.$test),
                  species1 = dplyr::first(.$species),
                  species2 = dplyr::last(.$species),
                  data = list(matrix(c(.$Absent,
                                       .$Present),
                                     ncol = 2)))) %>%
    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
    mutate(p.value = map_dbl(chi_test, "p.value"),
           dupl = duplicated(test)) %>%
    dplyr::filter(dupl == FALSE) %>%
    ungroup() %>%
    select(gene, species1, species2, p.value) %>%
    mutate(p.value = round(p.value, 5)) %>%
    dplyr::filter(p.value <= 0.05)
  return(df)
}

total_df <- acquired_report %>%
  left_join(mut_report, by = "id") %>%
  left_join(isolate_data[c("id", "species")], by = "id") %>%
  select(-nt_change)

mechanism_per_species <- total_df %>%
  select(id, species, everything()) %>%
  select(-id) %>%
  gather(key, value, -species) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  select(key, species, Present, Absent) %>%
  split(f = .$key)

intrinsic_names <- names(mut_report)
intrinsic_names <- intrinsic_names[-c(1,2)]

intrinsic_gene_perc <- mut_report %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, intrinsic_names) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  select(key, species, Percent) %>%
  dplyr::rename("gene" = key)

acquired_names <- names(acquired_report)
acquired_names <- acquired_names[-1]

acquired_genes_perc <- acquired_report %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, acquired_names) %>%
  group_by(key, value, species) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present / Total * 100, 1)) %>%
  select(key, species, Percent) %>%
  dplyr::rename("gene" = key)

total_perc <- rbind(intrinsic_gene_perc, acquired_genes_perc)

p_values_total <- suppressWarnings(lapply(mechanism_per_species, function(x) do_chisq_test(x))) %>%
  bind_rows() %>%
  left_join(total_perc, by = c("gene", c("species1" = "species"))) %>%
  dplyr::rename("Percent species 1" = Percent) %>%
  left_join(total_perc, by = c("gene", c("species2" = "species"))) %>%
  dplyr::rename("Percent species 2" = Percent) %>%
  dplyr::select(gene, species1, `Percent species 1`, species2, `Percent species 2`, p.value)

p_values_total

## Resistant isolates per antimicrobial per species
```{r}
percent_am_res <- mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
  select(species,contains("_R")) %>%
  gather(compound, value, -species) %>%
  mutate(compound = sub("_R", "", compound)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round((Present/Total)*100, 1))

percent_am_res
```

## Significant differences between the amount of resistant isolates
```{r}
cip_nal_res <- mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
  select(species,contains("_R")) %>%
  gather(compound, value, -species) %>%
  mutate(compound = sub("_R", "", compound)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  filter(compound %in% c("CIP","NAL")) %>%
  ungroup() %>%
  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
  spread(value, n, fill = 0) %>%
  select(compound, species, Present, Absent) %>%
  dplyr::rename("key" = compound) %>%
  split(f = .$key)

p_values_res <- suppressWarnings(lapply(cip_nal_res, function(x) do_chisq_test(x))) %>%
  bind_rows() %>%
  left_join(percent_am_res[,c("species","compound","Percent")], by = c("species1" = "species",
                                                                       "gene" = "compound")) %>%
  dplyr::rename("Percent_species1" = Percent) %>%
  left_join(percent_am_res[,c("species","compound","Percent")], by = c("species2" = "species",
                                                                       "gene" = "compound")) %>%
  dplyr::rename("Percent_species2" = Percent,
                "Compound" = gene) %>%
  select(Compound, species1, Percent_species1, species2, Percent_species2, p.value)

p_values_res

rest_res <- mic_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  mutate(SMX_R = ifelse(as.numeric(SMX) > 64, 1, 0),
         TMP_R = ifelse(as.numeric(TMP) > 2, 1, 0),
         CIP_R = ifelse(as.numeric(CIP) > 0.06, 1, 0),
         TET_R = ifelse(as.numeric(TET) > 8, 1, 0),
         NAL_R = ifelse(as.numeric(NAL) > 16, 1, 0),
         CTX_R = ifelse(as.numeric(CTX) > 0.25, 1, 0),
         CHL_R = ifelse(as.numeric(CHL) > 16, 1, 0),
         AMP_R = ifelse(as.numeric(AMP) > 8, 1, 0),
         GEN_R = ifelse(as.numeric(GEN) > 2, 1, 0)) %>%
  select(species,contains("_R")) %>%
  gather(compound, value, -species) %>%
  mutate(compound = sub("_R", "", compound)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  filter(compound %not_in% c("CIP","NAL")) %>%
  ungroup() %>%
  mutate(value = ifelse(value == 0, "Absent", "Present")) %>%
  spread(value, n, fill = 0) %>%
  select(compound, species, Present, Absent) %>%
  dplyr::rename("key" = compound) %>%
  split(f = .$key)
